---
title: "Absenteeism Analysis"
author: "Savaira Imran"
date: "2025-02-22"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# Load necessary libraries
knitr::opts_chunk$set(echo = TRUE)

# Read in the dataset
df <- read.csv("Absenteeism_at_work.csv", sep=";", header=TRUE)

# Check the structure of the data to understand the columns
str(df)
```

1.  Scatter Plot of Height vs. Weight

```{r}
# Plot scatter plot of height vs. weight (weight on x-axis)
plot(df$Weight, df$Height, main="Scatter Plot of Height vs. Weight",
     xlab="Weight (kg)", ylab="Height (cm)", pch=19, col=rgb(0.1, 0.2, 0.8, 0.5))
```

As weight goes up, the plot points become more scattered.

2.  Histogram

```{r}
# Plot histogram of hours of absences
hist(df$Absenteeism.time.in.hours, main="Histogram of Absenteeism Hours",
     xlab="Hours of Absence", col="lightblue", border="black", breaks=20)
```

Frequency goes down as hours of absence go up.

3.  Histogram of Age of Person Corresponding to each Absence

```{r}
# Plot histogram of age corresponding to each absence
hist(df$Age, main="Histogram of Age of Employee",
     xlab="Age", col="lightgreen", border="black", breaks=15)
```

Between age 35-40, the frequency is highest.

4.  Bar Plot of Hours by Month

```{r}
# Create a bar plot of total absentee hours by month
library(dplyr)
df_monthly <- df %>%
  group_by(Month.of.absence) %>%
  summarise(Total.absent.hours = sum(Absenteeism.time.in.hours, na.rm=TRUE))

barplot(df_monthly$Total.absent.hours, names.arg=df_monthly$Month.of.absence,
        main="Bar Plot of Absenteeism Hours by Month", xlab="Month", ylab="Total Absenteeism Hours",
        col="salmon", border="black")
```

Total hours are highest on month 3 and month 7.

5.  Box plots of hours by social smoker status

```{r}
# Box plot of absenteeism hours by smoker status
boxplot(Absenteeism.time.in.hours ~ Social.smoker, data=df, 
        main="Boxplot of Absenteeism Hours by Social Smoker",
        xlab="Social Smoker", ylab="Absenteeism Hours",
        col=c("lightblue", "pink"), border="black", notch=TRUE)
legend("topright", legend=c("Non-Smoker", "Smoker"), fill=c("lightblue", "pink"))
```

Non-smokers have a lot of outliers.

6.  Box Plots of Hours by Social Drinker Status

```{r}
# Box plot of absenteeism hours by social drinker status
boxplot(Absenteeism.time.in.hours ~ Social.drinker, data=df, 
        main="Boxplot of Absenteeism Hours by Social Drinker",
        xlab="Social Drinker", ylab="Absenteeism Hours",
        col=c("lightgreen", "lightcoral"), border="black", notch=TRUE)
legend("topright", legend=c("Non-Drinker", "Drinker"), fill=c("lightgreen", "lightcoral"))
```

The distribution of absenteeism hours for non-drinkers and drinkers are similar.
