---
title: "IMDB Movies Data Analysis"
author: "Savaira Imran"
date: "3/06/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Required Libraries

```{r}
library(ggplot2movies)
library(dplyr)
library(ggplot2)
library(tidyr)
```

## Load Data

```{r}
data(movies)
```

## 1. Range of Years of Production

```{r}
year_range <- range(movies$year)
year_range
```

## 2. Proportion of Movies with Budget Information

```{r}
budget_proportion <- mean(!is.na(movies$budget))
no_budget_proportion <- mean(is.na(movies$budget))

expensive_movies <- movies %>% 
  filter(!is.na(budget)) %>% 
  arrange(desc(budget)) %>% 
  select(title, year, budget) %>% 
  head(5)

budget_proportion
no_budget_proportion
expensive_movies
```

## 3. Top 5 Longest Movies

```{r}
longest_movies <- movies %>% 
  arrange(desc(length)) %>% 
  select(title, year, length) %>% 
  head(5)
longest_movies
```

## 4. Shortest and Longest Short Movies

```{r}
short_movies <- movies %>% filter(if_any(starts_with("Short"), ~ . == 1))
shortest_short <- short_movies %>% arrange(length) %>% select(title, length) %>% head(1)
longest_short <- short_movies %>% arrange(desc(length)) %>% select(title, length) %>% head(1)

shortest_short
longest_short
```

## 5. Number of Movies per Genre

```{r}
genre_counts <- movies %>% 
  summarise(
    Action = sum(Action),
    animation = sum(Animation),
    comedy = sum(Comedy),
    drama = sum(Drama),
    documentary = sum(Documentary),
    romance = sum(Romance),
    Short = sum(Short)
  ) %>% 
  pivot_longer(cols = everything(), names_to = "genre", values_to = "count")

ggplot(genre_counts, aes(x = genre, y = count, fill = genre)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Number of Movies per Genre", x = "Genre", y = "Count")
```

## 6. Average Rating by Genre

```{r}
ratings_per_genre <- movies %>% 
  summarise(
    Action = mean(rating[Action == 1], na.rm = TRUE),
    animation = mean(rating[Animation == 1], na.rm = TRUE),
    comedy = mean(rating[Comedy == 1], na.rm = TRUE),
    drama = mean(rating[Drama == 1], na.rm = TRUE),
    documentary = mean(rating[Documentary == 1], na.rm = TRUE),
    romance = mean(rating[Romance == 1], na.rm = TRUE),
    Short = mean(rating[Short == 1], na.rm = TRUE)
  ) %>% 
  pivot_longer(cols = everything(), names_to = "genre", values_to = "avg_rating")

ggplot(ratings_per_genre, aes(x = genre, y = avg_rating, fill = genre)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Average Rating per Genre", x = "Genre", y = "Average Rating")
```

## 7. Average Rating of Movies (2000-2005)

```{r}
avg_rating_2000_2005 <- movies %>% 
  filter(year >= 2000 & year <= 2005) %>%
  summarise(
    Action = mean(rating[Action == 1], na.rm = TRUE),
    animation = mean(rating[Animation == 1], na.rm = TRUE),
    comedy = mean(rating[Comedy == 1], na.rm = TRUE),
    drama = mean(rating[Drama == 1], na.rm = TRUE),
    documentary = mean(rating[Documentary == 1], na.rm = TRUE),
    romance = mean(rating[Romance == 1], na.rm = TRUE)
  ) %>% 
  pivot_longer(cols = everything(), names_to = "genre", values_to = "avg_rating")

ggplot(avg_rating_2000_2005, aes(x = genre, y = avg_rating, fill = genre)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Average Rating per Genre (2000-2005)", x = "Genre", y = "Average Rating")
```

## 8. Movie Production Trends (1990+)

```{r}
movies_1990 <- movies %>% 
  filter(year >= 1990) %>% 
  group_by(year) %>% 
  summarise(
    Action = sum(Action),
    animation = sum(Animation),
    comedy = sum(Comedy),
    drama = sum(Drama),
    documentary = sum(Documentary),
    romance = sum(Romance)
  ) %>% 
  pivot_longer(cols = -year, names_to = "genre", values_to = "count")

ggplot(movies_1990, aes(x = year, y = count, color = genre)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Movies Produced Per Year (1990+)", x = "Year", y = "Count")
```

## 9. Custom Questions

### a) Which genre has the highest-rated movie?

```{r}
highest_rated <- movies %>% 
  arrange(desc(rating)) %>% 
  select(title, rating) %>% 
  head(1)
highest_rated
```

### b) How does budget correlate with ratings?

```{r}
ggplot(movies, aes(x = budget, y = rating)) +
  geom_point(alpha = 0.3) +
  theme_minimal() +
  labs(title = "Budget vs. Rating", x = "Budget (USD)", y = "IMDB Rating")
```

### c) What is the distribution of movie lengths?

```{r}
ggplot(movies, aes(x = length)) +
  geom_histogram(binwidth = 10, fill = "blue", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Distribution of Movie Lengths", x = "Length (minutes)", y = "Count")
```

------------------------------------------------------------------------

This report provides insights into the IMDB movies dataset with visualizations and statistical summaries.
